### 2016-11-21 report/survey.tex
Added one platform and expanded explanation.
Added the bib file. Still need to make auto-generation of results,
and the same for hardware information and other stuff.

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{hyperref}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.12}
\usepackage{booktabs}
\usepackage{footnote}

\usepackage{fancyhdr, setspace}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
\lhead{}
\rhead{}
\usepackage{totcount}
\regtotcounter{page}
\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}

\usepackage{enumitem}
\setlist{parsep=0pt, listparindent=0.5cm}

\begin{document}
\begin{center}
{ \huge
	Simplifying Parallel Graph Processing: \\
}
{ \Large
	Survey of Existing Platforms \\
}
Sam Pollard (\href{mailto:spollard@cs.uoregon.edu}{spollard@cs.uoregon.edu}), University of Oregon \\
\today
\end{center}

This is a survey of existing graph analytics frameworks.

\section{Arya}
Below are the preliminary results by running two benchmarks on the research computer Arya.

Machine Specifications

% TODO: Get this to be filled out automatically wherever possible.

\begin{tabular}{lr}
CPU & 72 Core Intel Xeon E5 2699 v3\\ \hline
RAM & 256 GB DDR4 ? MHZ
\end{tabular}

Report Card

\begin{tabular}{lr}
Transport & \\ \hline
Network Topology & \\ \hline
Local Scheduling & \\ \hline
Runtime Feedback & \\ \hline
Approach & \\ \hline
Algorithmic Considerations & \\ \hline
\end{tabular}

Performance in millions of traversed edges per second (MTEPS)

\begin{tabular}{l|r|r|}
 & PowerGraph & OpenG \\ \hline
BFS & $87.4$ & $341$ \\ \hline
SSSP & $1.09$ & $3.08$ \\ \hline
\end{tabular}

\section{Graph Processing Taxonomy}
This is in the spirit of \cite{Doekemeijer:2015:GPFSurvey}. Here, ``|'' means ``or'' and ``+'' means ``and.'' FOSS means Free and Open Source Software. The quotes around ``yes'' for HPC mean that the product claims to be amenable to high performance computing. Whether these actually achieve their goal is one of the purposes of this project.
\begin{savenotes}
	\begin{table}[htb]
		\centering
		% \small
		\pgfplotstabletypeset[
			col sep=comma,
			string type, % Makes the .style={string type} redundant
			every head row/.style={after row=\midrule},
			columns/Name/.style={string type, column type={l|}},
			columns/Type/.style={string type},
			columns/HPC/.style={string type},
			columns/Parallelism/.style={string type},
			columns/Target/.style={string type},
			columns/FOSS/.style={string type},
			columns/Source/.style={string type},
			columns/Notes/.style={
				preproc cell content/.style={@cell content=
					\ifx&##1&% Only make a footnote if the cell is nonempty
						##1
					\else
						\footnote{##1}
					\fi}
			},
		]{platforms.csv}
	\end{table}
\end{savenotes}

\bibliographystyle{acm}
\bibliography{drp}

\end{document}