### 2016-12-14 taxonomy/taxonomy.tex
Added a template for graph processing taxonomy.
This will eventually describe the current state of the art.

\documentclass[11pt]{scrartcl}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{hyperref}
\usepackage{pgfplotstable, booktabs}
\pgfplotsset{compat=1.12}
\usepackage{datetime}

\usepackage{fancyhdr, setspace}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\setlength{\headheight}{15pt} % Removes warnings. Set to 30pt if 2 lines on the header.
\lhead{}
\rhead{}
\usepackage{totcount}
\regtotcounter{page}
\cfoot{\ifnum\totvalue{page} > 1 \thepage \else\fi}

\begin{document}
\title{The State of Graph Processing}
\subtitle{APIs, Libraries, Platforms, and Programming Languages}
\author{Samuel Pollard}
\date{\today}
\maketitle

\section{Introduction}

\end{document}